The unphysicality of Hilbert spaces

### 2024-01-23 Articles/2023-HilbertQuantum/HilbertQuantum.tex
Add \usepackage{breakurl} so bibliography formats correctly on arxiv
HilbertQuantum: removing "only space"

\usepackage{assumptionsofphysics}
\usepackage{tikz}

@@

\begin{abstract}
We show that Hilbert spaces should not be considered the ``correct'' spaces to represent quantum states mathematically. We first prove that the requirements posited by complex inner product spaces are physically justified. We then show that completeness in the infinite-dimensional case requires the inclusion of states with infinite expectations, coordinate transformations that take finite expectations to infinite ones and vice-versa, and time evolutions that transform finite expectations to infinite ones in finite time. This makes Hilbert spaces physically unsound as they model a potential infinity as an actual infinity. We suspect that at least some problems in quantum theory related to infinities may be ultimately caused by the wrong space being used. We strongly believe a better solution can be found, and we look at Schwartz spaces for inspiration, as, among other things, they guarantee that the expectation of all polynomials of position and momentum are finite, their elements are uniquely identified by these expectations, and they are the only space closed under Fourier transform.
\end{abstract}

@@

In this work we briefly present our findings, leaving more mathematical detail to the appendix. We will see that complex inner product spaces are indeed needed, and should be considered physical. We will see that the problem lies in requiring completeness, which should be understood as an unphysical mathematical requirement. We will show that other options exist, offering Schwartz spaces as an example. These guarantee the expectation values of all polynomials of position and momentum to be finite, and they are also the only space closed under Fourier transform.

@@

\begin{itemize}
\item its elements are fully identified by the expectations - two Schwartz functions are the same if and only if all expectations are the same; quantum tomography on all polynomials fully reconstructs the state
\item it is an inner product space with the same norm as $L^2$
\item it is a dense subset of $L^2$ - any element of the Hilbert space can be approximated by a Schwartz function with an arbitrary level of precision
\item it is complete with respect to the expectation values - a sequence of functions for which all expectation values of all polynomials of position and momentum converge will converge in the Schwartz space
\item its closure over all Cauchy sequences recovers the Hilbert space - the only thing missing are the Cauchy sequences for which the limit of at least one expectation is not well defined
\item it is the only space closed under Fourier transform - the Fourier transform of a Schwartz function is a Schwartz function
\item it plays a fundamental role in the theory of distributions - objects like the $\delta$-function are mathematically defined on top of the Schwartz space
\item the Hermite polynomials, the solutions of the harmonic oscillator, are Schwartz functions
\end{itemize}
These properties make the Schwartz space a much more reasonable candidate to capture the physics.\cite{albert2022bosonic} The convergence of expectation values of actual observables replaces convergence over vectors in an abstract space.\footnote{Compare this statement with the ``weak convergence'' of von Neumann algebras.} To again be clear, we are not proposing this as a solution. We are merely pointing out that alternatives do exist and should be fully explored.\footnote{Note that $C^*$-algebras model observables and therefore do not solve the problem at hand of modeling states.}

@@

The Schwartz space also has a very important property: it is the only space that is closed under the Fourier transform. That is, the Fourier transform of a Schwartz function is a Schwartz function, and it is the only set of functions with this property. Physically, this means that we can always express a wavefunction in either the position or the momentum basis. Mathematically, the theory of tempered distributions, which allows us to define objects like the Dirac delta distribution, is built on top of the Schwartz space. Therefore, this space already plays a fundamental role in the definitions of the tools we already use in physics.

>>>>

\usepackage{assumptionsofphysics}
\usepackage{tikz}
\usepackage{breakurl}

@@

\begin{abstract}
We show that Hilbert spaces should not be considered the ``correct'' spaces to represent quantum states mathematically. We first prove that the requirements posited by complex inner product spaces are physically justified. We then show that completeness in the infinite-dimensional case requires the inclusion of states with infinite expectations, coordinate transformations that take finite expectations to infinite ones and vice-versa, and time evolutions that transform finite expectations to infinite ones in finite time. This makes Hilbert spaces physically unsound as they model a potential infinity as an actual infinity. We suspect that at least some problems in quantum theory related to infinities may be ultimately caused by the wrong space being used. We strongly believe a better solution can be found, and we look at Schwartz spaces for inspiration, as, among other things, they guarantee that the expectation of all polynomials of position and momentum are finite, their elements are uniquely identified by these expectations, and they are closed under Fourier transform.
\end{abstract}

@@

In this work we briefly present our findings, leaving more mathematical detail to the appendix. We will see that complex inner product spaces are indeed needed, and should be considered physical. We will see that the problem lies in requiring completeness, which should be understood as an unphysical mathematical requirement. We will show that other options exist, offering Schwartz spaces as an example. These guarantee the expectation values of all polynomials of position and momentum to be finite, and they are also closed under Fourier transform.

@@

\begin{itemize}
\item its elements are fully identified by the expectations - two Schwartz functions are the same if and only if all expectations are the same; quantum tomography on all polynomials fully reconstructs the state
\item it is an inner product space with the same norm as $L^2$
\item it is a dense subset of $L^2$ - any element of the Hilbert space can be approximated by a Schwartz function with an arbitrary level of precision
\item it is complete with respect to the expectation values - a sequence of functions for which all expectation values of all polynomials of position and momentum converge will converge in the Schwartz space
\item its closure over all Cauchy sequences recovers the Hilbert space - the only thing missing are the Cauchy sequences for which the limit of at least one expectation is not well defined
\item it is closed under Fourier transform - the Fourier transform of a Schwartz function is a Schwartz function
\item it plays a fundamental role in the theory of distributions - objects like the $\delta$-function are mathematically defined on top of the Schwartz space
\item the Hermite polynomials, the solutions of the harmonic oscillator, are Schwartz functions
\end{itemize}
These properties make the Schwartz space a much more reasonable candidate to capture the physics.\cite{albert2022bosonic} The convergence of expectation values of actual observables replaces convergence over vectors in an abstract space.\footnote{Compare this statement with the ``weak convergence'' of von Neumann algebras.} To again be clear, we are not proposing this as a solution. We are merely pointing out that alternatives do exist and should be fully explored.\footnote{Note that $C^*$-algebras model observables and therefore do not solve the problem at hand of modeling states.}

@@

The Schwartz space also has another important property: it is closed under the Fourier transform. That is, the Fourier transform of a Schwartz function is a Schwartz function. Physically, this means that we can always express a wavefunction in either the position or the momentum basis. Mathematically, the theory of tempered distributions, which allows us to define objects like the Dirac delta distribution, is built on top of the Schwartz space. Therefore, this space already plays a fundamental role in the definitions of the tools we already use in physics.
